\documentclass{article}
\usepackage{fuzz}

\begin{document}

%\begin{zsection}	 \SECTION ESELSystem1 \parents~standard\_toolkit
%\end{zsection}

\begin{axdef}
	MAX\_ESEL : \nat \\
 MAX\_PID : \nat 
\where
 MAX\_ESEL = 3 \\
 MAX\_PID = 2
\end{axdef}

\begin{zed}
	ESID ::= ES \ldata 1 \upto MAX\_ESEL \rdata
\end{zed}

\begin{zed}
	PID ::= PD \ldata 1 \upto MAX\_PID \rdata
\end{zed}

\begin{zed}
	Price == \nat
\end{zed}

\begin{zed}
	UStatus ::= uok | ufail
\end{zed}

\begin{zed}
	FStatus ::= fail \ldata ESID \rdata | NA
\end{zed}

\begin{zed}
	Controller\_State \defs [~  Controller\_pumap : ESID \pfun PID ; Controller\_ppmap : PID \pfun Price ; Controller\_response : PID \pfun ( \power FStatus )  ~]
\end{zed}

\begin{zed}
	ESEL\_ES1\_State \defs [~  ESEL\_ES1\_price : Price ; ESEL\_ES1\_status : UStatus  ~]
\end{zed}

\begin{zed}
	ESEL\_ES3\_State \defs [~  ESEL\_ES3\_price : Price ; ESEL\_ES3\_status : UStatus  ~]
\end{zed}

\begin{zed}
	ESEL\_ES2\_State \defs [~  ESEL\_ES2\_price : Price ; ESEL\_ES2\_status : UStatus  ~]
\end{zed}

\begin{zed}
	State \defs Controller\_State \land ESEL\_ES1\_State \land ESEL\_ES3\_State \land ESEL\_ES2\_State
\end{zed}

\begin{zed}
	Init \defs [~  State' | Controller\_pumap' =~\emptyset \land Controller\_ppmap' =~\emptyset \land Controller\_response' =~\emptyset \land ESEL\_ES1\_price' = 0 \land ESEL\_ES1\_status' = uok \land ESEL\_ES3\_price' = 0 \land ESEL\_ES3\_status' = uok \land ESEL\_ES2\_price' = 0 \land ESEL\_ES2\_status' = uok  ~]
\end{zed}

\begin{zed}
	Controller\_Init \defs [~  Controller\_State' ; Controller\_pumap : ESID \pfun PID ; Controller\_response : PID \pfun ( \power FStatus ) ; Controller\_ppmap : PID \pfun Price ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | Controller\_pumap' =~\emptyset \land Controller\_ppmap' =~\emptyset \land Controller\_response' =~\emptyset  ~]
\end{zed}

\begin{zed}
	Controller\_UpdateMap \defs [~  \Delta Controller\_State ; map? : ESID \pfun PID ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | Controller\_pumap' = Controller\_pumap \oplus map? \land Controller\_ppmap' = Controller\_ppmap \land Controller\_response' = Controller\_response  ~]
\end{zed}

\begin{zed}
	Controller\_UpdateAllMap \defs [~  \Delta Controller\_State ; map? : ESID \pfun PID ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | Controller\_pumap' = map? \land Controller\_ppmap' = Controller\_ppmap \land Controller\_response' = Controller\_response  ~]
\end{zed}

\begin{zed}
	Controller\_NewPrice \defs [~  \Delta Controller\_State ; price? : PID \pfun Price ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | Controller\_ppmap' = Controller\_ppmap \oplus price? \land Controller\_pumap' = Controller\_pumap \land Controller\_response' = Controller\_response  ~]
\end{zed}

\begin{zed}
	Controller\_AllNewPrice \defs [~  \Delta Controller\_State ; price? : PID \pfun Price ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | Controller\_ppmap' = price? \land Controller\_pumap' = Controller\_pumap \land Controller\_response' = Controller\_response  ~]
\end{zed}

\begin{zed}
	Controller\_AddOneFailure \defs [~  \Delta Controller\_State ; pid? : PID ; fst? : FStatus ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | ( pid? \in \dom Controller\_response \implies Controller\_response' = Controller\_response \oplus \{ pid? \mapsto ( Controller\_response ( pid? ) \cup \{ fst? \} ) \} ) \land ( pid? \notin \dom Controller\_response \implies Controller\_response' = Controller\_response \cup \{ pid? \mapsto \{ fst? \} \} ) \land Controller\_ppmap' = Controller\_ppmap \land Controller\_pumap' = Controller\_pumap  ~]
\end{zed}

\begin{zed}
	Controller\_OP\_pumap \defs [~  \Xi Controller\_State ; Controller\_pumap! : ESID \pfun PID ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | Controller\_pumap! = Controller\_pumap  ~]
\end{zed}

\begin{zed}
	Controller\_OP\_response \defs [~  \Xi Controller\_State ; Controller\_response! : PID \pfun ( \power FStatus ) ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | Controller\_response! = Controller\_response  ~]
\end{zed}

\begin{zed}
	Controller\_OP\_ppmap \defs [~  \Xi Controller\_State ; Controller\_ppmap! : PID \pfun Price ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | Controller\_ppmap! = Controller\_ppmap  ~]
\end{zed}

\begin{zed}
	Controller\_Init\_fOp \defs [~  \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | \lnot \pre Controller\_Init  ~]
\end{zed}

\begin{zed}
	Controller\_UpdateMap\_fOp \defs [~  \Xi Controller\_State ; map? : ESID \pfun PID ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | \lnot \pre Controller\_UpdateMap  ~]
\end{zed}

\begin{zed}
	Controller\_UpdateAllMap\_fOp \defs [~  \Xi Controller\_State ; map? : ESID \pfun PID ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | \lnot \pre Controller\_UpdateAllMap  ~]
\end{zed}

\begin{zed}
	Controller\_NewPrice\_fOp \defs [~  \Xi Controller\_State ; price? : PID \pfun Price ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | \lnot \pre Controller\_NewPrice  ~]
\end{zed}

\begin{zed}
	Controller\_AllNewPrice\_fOp \defs [~  \Xi Controller\_State ; price? : PID \pfun Price ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | \lnot \pre Controller\_AllNewPrice  ~]
\end{zed}

\begin{zed}
	Controller\_AssgnOp0 \defs [~  \Delta Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | Controller\_response' =~\emptyset \land Controller\_ppmap' = Controller\_ppmap \land Controller\_pumap' = Controller\_pumap  ~]
\end{zed}

\begin{zed}
	Controller\_AddOneFailure\_fOp \defs [~  \Xi Controller\_State ; pid? : PID ; fst? : FStatus ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | \lnot \pre Controller\_AddOneFailure  ~]
\end{zed}

\begin{zed}
	ESEL\_ES1\_Init \defs [~  ESEL\_ES1\_State' ; ESEL\_ES1\_status : UStatus ; ESEL\_ES1\_price : Price ; \Xi Controller\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | ESEL\_ES1\_price' = 0 \land ESEL\_ES1\_status' = uok  ~]
\end{zed}

\begin{zed}
	ESEL\_ES1\_OP\_status \defs [~  \Xi ESEL\_ES1\_State ; ESEL\_ES1\_status! : UStatus ; \Xi Controller\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | ESEL\_ES1\_status! = ESEL\_ES1\_status  ~]
\end{zed}

\begin{zed}
	ESEL\_ES1\_OP\_price \defs [~  \Xi ESEL\_ES1\_State ; ESEL\_ES1\_price! : Price ; \Xi Controller\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | ESEL\_ES1\_price! = ESEL\_ES1\_price  ~]
\end{zed}

\begin{zed}
	ESEL\_ES1\_Init\_fOp \defs [~  \Xi ESEL\_ES1\_State ; \Xi Controller\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | \lnot \pre ESEL\_ES1\_Init  ~]
\end{zed}

\begin{zed}
	ESEL\_ES1\_AssgnOp1 \defs [~  \Delta ESEL\_ES1\_State ; x? : Price ; \Xi Controller\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | ESEL\_ES1\_price' = x? \land ESEL\_ES1\_status' = ESEL\_ES1\_status  ~]
\end{zed}

\begin{zed}
	ESEL\_ES1\_AssgnOp2 \defs [~  \Delta ESEL\_ES1\_State ; \Xi Controller\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | ESEL\_ES1\_status' = uok \land ESEL\_ES1\_price' = ESEL\_ES1\_price  ~]
\end{zed}

\begin{zed}
	ESEL\_ES1\_AssgnOp3 \defs [~  \Delta ESEL\_ES1\_State ; \Xi Controller\_State ; \Xi ESEL\_ES3\_State ; \Xi ESEL\_ES2\_State | ESEL\_ES1\_status' = ufail \land ESEL\_ES1\_price' = ESEL\_ES1\_price  ~]
\end{zed}

\begin{zed}
	ESEL\_ES3\_Init \defs [~  ESEL\_ES3\_State' ; ESEL\_ES3\_price : Price ; ESEL\_ES3\_status : UStatus ; \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES2\_State | ESEL\_ES3\_price' = 0 \land ESEL\_ES3\_status' = uok  ~]
\end{zed}

\begin{zed}
	ESEL\_ES3\_OP\_price \defs [~  \Xi ESEL\_ES3\_State ; ESEL\_ES3\_price! : Price ; \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES2\_State | ESEL\_ES3\_price! = ESEL\_ES3\_price  ~]
\end{zed}

\begin{zed}
	ESEL\_ES3\_OP\_status \defs [~  \Xi ESEL\_ES3\_State ; ESEL\_ES3\_status! : UStatus ; \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES2\_State | ESEL\_ES3\_status! = ESEL\_ES3\_status  ~]
\end{zed}

\begin{zed}
	ESEL\_ES3\_Init\_fOp \defs [~  \Xi ESEL\_ES3\_State ; \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES2\_State | \lnot \pre ESEL\_ES3\_Init  ~]
\end{zed}

\begin{zed}
	ESEL\_ES3\_AssgnOp4 \defs [~  \Delta ESEL\_ES3\_State ; x? : Price ; \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES2\_State | ESEL\_ES3\_price' = x? \land ESEL\_ES3\_status' = ESEL\_ES3\_status  ~]
\end{zed}

\begin{zed}
	ESEL\_ES3\_AssgnOp5 \defs [~  \Delta ESEL\_ES3\_State ; \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES2\_State | ESEL\_ES3\_status' = uok \land ESEL\_ES3\_price' = ESEL\_ES3\_price  ~]
\end{zed}

\begin{zed}
	ESEL\_ES3\_AssgnOp6 \defs [~  \Delta ESEL\_ES3\_State ; \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES2\_State | ESEL\_ES3\_status' = ufail \land ESEL\_ES3\_price' = ESEL\_ES3\_price  ~]
\end{zed}

\begin{zed}
	ESEL\_ES2\_Init \defs [~  ESEL\_ES2\_State' ; ESEL\_ES2\_price : Price ; ESEL\_ES2\_status : UStatus ; \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State | ESEL\_ES2\_price' = 0 \land ESEL\_ES2\_status' = uok  ~]
\end{zed}

\begin{zed}
	ESEL\_ES2\_OP\_price \defs [~  \Xi ESEL\_ES2\_State ; ESEL\_ES2\_price! : Price ; \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State | ESEL\_ES2\_price! = ESEL\_ES2\_price  ~]
\end{zed}

\begin{zed}
	ESEL\_ES2\_OP\_status \defs [~  \Xi ESEL\_ES2\_State ; ESEL\_ES2\_status! : UStatus ; \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State | ESEL\_ES2\_status! = ESEL\_ES2\_status  ~]
\end{zed}

\begin{zed}
	ESEL\_ES2\_Init\_fOp \defs [~  \Xi ESEL\_ES2\_State ; \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State | \lnot \pre ESEL\_ES2\_Init  ~]
\end{zed}

\begin{zed}
	ESEL\_ES2\_AssgnOp7 \defs [~  \Delta ESEL\_ES2\_State ; x? : Price ; \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State | ESEL\_ES2\_price' = x? \land ESEL\_ES2\_status' = ESEL\_ES2\_status  ~]
\end{zed}

\begin{zed}
	ESEL\_ES2\_AssgnOp8 \defs [~  \Delta ESEL\_ES2\_State ; \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State | ESEL\_ES2\_status' = uok \land ESEL\_ES2\_price' = ESEL\_ES2\_price  ~]
\end{zed}

\begin{zed}
	ESEL\_ES2\_AssgnOp9 \defs [~  \Delta ESEL\_ES2\_State ; \Xi Controller\_State ; \Xi ESEL\_ES1\_State ; \Xi ESEL\_ES3\_State | ESEL\_ES2\_status' = ufail \land ESEL\_ES2\_price' = ESEL\_ES2\_price  ~]
\end{zed}

\end{document}
