\section{Requirements Modelling} %Sec 4 %=================================

 Every requirement is modelled as a process over hardware and TSP events,
 with some requirement-specific marker events added in.

 Most requirement models encode the relevant satisfaction property in terms
 of either the presence or absence of deadlock, as appropriate.



% include "Platform.csp"

\section{Initialisation Management} %------------------------- -}

\subsection{PK-79}
The partitioning kernel shall start its execution by initialising the system according to the defined configuration.

\subsection{PK-84}
After PK initialization phase, the partitioning kernel shall start the execution of the default scheduling plan as defined in configuration.

\subsection{PK-86}
The partitioning Kernel configuration shall allow defining the entry start address of each partition, this entry start address being in RAM or in NVR (execute in place partition).

\subsection{PK-101}
Before restarting a partition from its entry start address, the partitioning kernel shall initialize its context.

\section{Partition Mode Management} % -------------------------

Things we need to observe:
\begin{itemize}
  \item what code (kernel, partition1, ...) is currently running
  \item what priviledge mode (user, supervisor) the processor is currently in
  \item which code segment (initialisation, main loop, ...) of an entity is running, and when it starts and finishes.
  \item we need to observe when the kernel starts a partition (cold or warm).
\end{itemize}

\subsection{PK-230}
Partitions shall be executed in user mode. (Req 5)

We track all ftch events and processor mode changes

\begin{circus}
\circchannel pk230fail\\

\circprocess PK230 \circdef \circbegin\\

PK230sup
 \circdef ((ftch.KNL?i?op \then PK230sup)
    \extchoice
    (ftch.P1?i?op \then pk230fail \then \Stop)
    \extchoice
    (ftch.P1?i?op \then pk230fail \then \Stop)
    \extchoice
    (pm.USR \then PK230usr)
    \extchoice
    (pm.SUP \then PK230sup))\\

PK230usr
 \circdef ((ftch?ent?i?op \then PK230usr)
    \extchoice
    (pm.USR \then PK230usr)
    \extchoice
    (pm.SUP \then PK230sup))\\
\circspot  PK230sup % assumes supervisor mode when started
\circend
\end{circus}
we now give two tests: good behaviour and bad behaviour

\begin{assert}
"PK230GOODCODE = ftch.KNL.NOP.0 -> ftch.KNL.S0.0 -> ftch.P1.NOP.0 -> ftch.P1.NOP.1 -> ftch.P1.NOP.0 -> ftch.P1.NOP.1 -> ftch.KNL.S1.0 -> ftch.KNL.NOP.1 -> PK230GOODCODE"\\


\also "assert PK230GOODCODE :[divergence free]"\\
\also "assert PK230GOODCODE :[deadlock free[F]]"\\

\also "PK230GOODSYS = PMM [| {| ftch |} |] PK230GOODCODE"\\

\also "assert PK230GOODSYS :[divergence free]"\\
\also "assert PK230GOODSYS :[deadlock free[F]]"\\

\also "PK230GOOD = PK230 [| {| ftch, pm |} |] PK230GOODSYS"\\

\also "assert PK230GOOD :[divergence free]"\\
\also "assert PK230GOOD :[deadlock free[F]]"\\
\end{assert}

\begin{assert}
"PK230BADCODE = ftch.KNL.NOP.0 -> ftch.KNL.S1.0 -> ftch.P1.NOP.0 -> ftch.P1.NOP.1 -> ftch.P1.NOP.0 -> ftch.P1.NOP.1 -> ftch.KNL.S1.0 -> ftch.KNL.NOP.1 -> PK230BADCODE"\\

   \also "assert PK230BADCODE :[divergence free]"\\
   \also "assert PK230BADCODE :[deadlock free[F]]"\\

   \also "PK230BADSYS = PMM [| {| ftch |} |] PK230BADCODE"\\

   \also "assert PK230BADSYS :[divergence free]"\\
   \also "assert PK230BADSYS :[deadlock free[F]]"\\

   \also "PK230BAD = PK230 [| {| ftch, pm |} |] PK230BADSYS"\\

   \also "assert PK230BAD :[divergence free]"\\
   \also "assert PK230BAD :[deadlock free[F]]"
\end{assert}
\subsection{PK-98}
When a partition enters in the ACTIVE mode, the Partitioning Kernel shall start executing the partition at the entry start address as defined in configuration.

\subsection{PK-541}
When a partition enters in the ACTIVE mode,  its associated vCPU shall have been reset.

\subsection{PK-97}
The starting type of a partition shall be WARM or COLD. (Req 8)

We track partition start events, from both kernel and partition perspective

\begin{circus}
\circchannel pk97fail
\end{circus}


\subsection{PK-100}
The partitioning kernel shall allow an authorized partition to start any partition (STOPED) mode to ACTIVE mode)  with a WARM or COLD starting type. (Req 9)

%{- 4.3.... not yet in here... --------------------------- -}
